---
title: "ThaiGov's 2022 Budget+"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", dpi = 320)
Sys.setlocale("LC_ALL", "Thai")
options(encoding = "UTF-8")
# set.seed(112)
library(dplyr)
library(skimr)
library(ggplot2)
library(stringr)
library(forcats)
library(scales)
library(targets)
library(rvest)
```

```{r, include = FALSE}
library(targets)
```

```{r, include = FALSE}
tar_unscript()
```

![<https://img.shields.io/badge/open%20data-thailand-blue>](https://img.shields.io/badge/open%20data-thailand-blue)

This repository contains a richer version of [2022 Thai Government Budget data](https://github.com/kaogeek/thailand-budget-pdf2csv).

Improvements:

-   Translated most of the character columns from Thai to English using [Google Translation API](https://cloud.google.com/translate).
-   Added address fields extracted from `item_description`.

The data can be viewed and downloaded from here:

> <https://docs.google.com/spreadsheets/d/1rKR1kLuSDssT0_xLpGE_oRm2tPD5ZRhzErWq-8UzH6A/edit?usp=sharing>

So far I have only translated `ministry`, `budgetary_unit`, `budget_plan`, `output`, `project`, `category_lv1`, `category_lv2`, and `category_lv3` columns using my free Google monthly quota. If you are interested to contribute, please submit a pull request with other columns translated to English. Feel free to use the R code below. :)

```{r echo = FALSE, warning = FALSE, message = FALSE}
set.seed(112)
tar_load(budget_plus)
cols_to_select <- names(budget_plus)[grepl("_en", names(budget_plus))] %>%
  c(gsub("_en", "", .), .) %>%
  .[!grepl("province|district|subdistrict|region", .)] %>%
  sort()
  
budget_plus %>%
    dplyr::select(cols_to_select, starts_with(c("province", "district", "subdistrict", "region"))) %>% 
    dplyr::slice_sample(n = 2) %>%
    knitr::kable()
```

```{r total-budget-by-ministry, echo = FALSE, fig.height=10}
billion_format <- unit_format(unit = "B", scale = 1e-9, big.mark = ",")

plot_data <-
  tar_read(budget_plus) %>%
  mutate(ministry_en = ifelse(grepl("Ministry of Higher Education", ministry_en) == TRUE,
    "Ministry of Higher Education, Science, Research and Innovation",
    ministry_en
  )) %>%
  mutate(ministry_en = ifelse(grepl("^Provinces", ministry_en) == TRUE,
    "Provinces and Provincial Groups",
    ministry_en
  )) %>%
  group_by(ministry_en, region_en) %>%
  summarise(amount = sum(amount, na.rm = TRUE))

plot_data %>%
  group_by(ministry_en) %>%
  summarise(amount = sum(amount, na.rm = TRUE)) %>%
  ggplot(data = ., aes(
  x = amount,
  y = forcats::fct_reorder(stringr::str_wrap(ministry_en, 60), amount)
)) +
  geom_col(fill = "#4b5320") +
  geom_text(aes(label = billion_format(amount)), hjust = -0.2, size = 3) +
  scale_x_continuous(labels = billion_format, limits = c(0, max(plot_data$amount, na.rm = TRUE) * 1.2)) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Amount in Billion Thai Bahts (พันล้าน)", y = "Ministry", title = "Total Amount of Budget by Ministry")
```

```{r total-budget-by-ministry-and-region, echo = FALSE, fig.height=10}
plot_data %>%
  dplyr::filter(!is.na(region_en)) %>%
  ggplot(data = ., aes(
    x = amount,
    y = forcats::fct_reorder(stringr::str_wrap(ministry_en, 60), amount),
    fill = region_en
  )) +
  geom_col() +
  scale_x_continuous(labels = billion_format) +
  scale_fill_brewer(palette = "Dark2") +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Amount in Billion Thai Bahts (พันล้าน)", y = "Ministry", title = "Total Amount of Budget by Ministry and Region", fill = "Region", caption = "*Not properly sorted by total amount")
```

# Targets

```{targets, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("dplyr", "ggplot2", "googlesheets4", "tidyr", "skimr", "magrittr", "googleLanguageR", "data.table"))
merge_translation <- function(x, translation, column) {
  translation %>%
    merge(x, ., by.x = column, by.y = "text", all.x = TRUE) %>%
    select(-detectedSourceLanguage) %>%
    rename_with( ~ gsub("translatedText", paste0(column, "_en"), .x, fixed = TRUE))
}
```

```{targets budget_raw, tar_simple = TRUE}
read_sheet("https://docs.google.com/spreadsheets/d/1yyWXSTbq3CD_gNxks-krcSBzbszv3c_2Nq54lckoQ24/edit#gid=1625073248")
```

```{targets data-prep}
list(
  tar_target(budget,
    budget_raw %>%
      janitor::clean_names()
  ),
  tar_target(unique_sentences, {
    budget %>%
      select(
        ministry,
        budgetary_unit,
        budget_plan,
        output,
        project,
        starts_with("category"),
        item_description
      ) %>% {
        as.character(unique(unlist(.)))
      }
  })
)
```

```{targets translation}
list(
  tar_target(translated_ministry, {
    budget$ministry %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_budgetary_unit, {
    budget$budgetary_unit %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_budget_plan, {
    budget$budget_plan %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_project, {
    budget$project %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_output, {
    budget$output %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_category_lv1, {
    budget$category_lv1 %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_category_lv2, {
    budget$category_lv2 %>%
      unique() %>%
      gl_translate(target = "en")
  }),
  tar_target(translated_category_lv3, {
    budget$category_lv2 %>%
      unique() %>%
      gl_translate(target = "en")
  })
)
```

```{targets budget_plus, tar_simple = TRUE}
budget %>%
  merge_translation(translated_ministry, "ministry") %>%
  merge_translation(translated_budgetary_unit, "budgetary_unit") %>%
  merge_translation(translated_budget_plan, "budget_plan") %>%
  merge_translation(translated_output, "output") %>%
  merge_translation(translated_project, "project") %>%
  merge_translation(translated_category_lv1, "category_lv1") %>%
  merge_translation(translated_category_lv2, "category_lv2") %>%
  merge_translation(translated_category_lv3, "category_lv3") %>% 
  dplyr::mutate(province = stringr::str_extract(item_description, "จังหวัด[^\\s]+|กรุงเทพมหานคร")) %>%
  dplyr::mutate(district = stringr::str_extract(item_description, "อำเภอ[^\\s]+|เขต[^\\s]+")) %>%
  dplyr::mutate(subdistrict = stringr::str_extract(item_description, "ตำบล[^\\s]+|แขวง[^\\s]+")) %>%
  dplyr::mutate(province = gsub("จังหวัด", "", province)) %>%
  merge(dplyr::select(provinces, starts_with("district")), by = "district", all.x = TRUE) %>%
  merge(dplyr::select(distinct(provinces, province, .keep_all = T), starts_with("province"), region_en), by = "province", all.x = TRUE) %>%
  select(names(budget), everything())
```

```{targets extract-addresses}
list(tar_target(provinces, {
  exclude_bkk <-
    rvest::read_html("https://en.wikipedia.org/wiki/List_of_districts_of_Thailand") %>%
    rvest::html_element(css = "table") %>%
    rvest::html_table() %>%
    janitor::clean_names() %>%
    setNames(c("district_en", "district", "province_en", "province", "region_en")) %>%
    dplyr::mutate(postcode = NA_integer_) %>%
    dplyr::mutate(district = paste0("อำเภอ", district))
  bkk <-
    rvest::read_html("https://en.wikipedia.org/wiki/List_of_districts_of_Bangkok") %>%
    rvest::html_element(css = "table") %>%
    rvest::html_table() %>%
    janitor::clean_names() %>%
    dplyr::mutate(province_en = "Bangkok", province = "กรุงเทพ", region_en = "Centre") %>%
    dplyr::select(district_en = district_khet, district = thai, province_en, province, region_en, postcode = post_code) %>%
    dplyr::mutate(district = paste0("เขต", district))
  rbind(bkk,exclude_bkk)
}))
```

```{targets upload_to_gsheet, tar_simple = TRUE}
# googlesheets4::gs4_create("65_thailand_budget_extracted_b4_cleansing_with_ENlang",
#            sheets = list(DATA = head(translated_budget, 10)))
googlesheets4::sheet_write(
  data = budget_plus,
  ss = "https://docs.google.com/spreadsheets/d/1rKR1kLuSDssT0_xLpGE_oRm2tPD5ZRhzErWq-8UzH6A/edit?usp=sharing",
  sheet = "DATA"
)
```

# Pipeline

```{r}
tar_make()
```

```{r target-visnetwork, out.width="100%"}
tar_visnetwork()
```

# Data summary

``` r
library(targets)
library(skimr)
```

```{r}
tar_read(budget_plus) %>%
  skimr::skim()
```
